### 2025-02-28 paper/EDA.qmd
Write up started

---
title: "Effects of Different Pain Relievers and Dosages on Cognitive Retention"
subtitle: "Exploratory Data Analysis (EDA)"
author: 
  - Navya Hooda
  - Mohammed Yusuf Shaikh
  - Shivank Goel
  - Vanshika Vanshika
  - Jena Shah
date: today
date-format: long
number-sections: true
format: pdf
---

# Introduction

## Relevant Background
Pain can affect how well we perform everyday tasks in situations where focus and mental clarity
are crucial. Aspirin, Paracetamol, and Tramadol are pain relievers used commonly to reduce
physical discomfort, but we don’t know much about their impacts on cognitive functions. Tasks
requiring mental effort, such as recalling information, are harder to perform while in pain.
This study is interested in how different types of common pain relievers at standard dosages
(Aspirin, Paracetamol, and Tramadol) may affect outcomes of tasks involving memory tests. We
also analyze the effects of different dosages of each drug on memory function. The concept of
recalling information is used in various settings, so understanding how pain relief can impact
performance can be valuable to help with pain management without an impact on productivity.

## Research Questions
1. *How does the type of pain reliever (Aspirin 500 mg, Paracetamol 500 mg, Tramadol 50 mg, and Placebo) affect cognitive task performance?*
2. *How does the dosage (low vs. high) of each drug affect cognitive retention, while accounting for confounding factors like age?*

## Study Design  
This study involves *7 treatment groups*, with **30 participants** assigned to each. The groups are as follows:

1. Aspirin 500 mg
2. Aspirin 1000 mg
3. Paracetamol 500 mg
4. Paracetamol 1000 mg
5. Tramadol 50 mg
6. Tramadol 100 mg
7. Placebo

This results in **210 total observations** (7 × 30 = 210).

## Confounding Variable: Age  
To reduce variability, we consider **age** as a **confounding variable**. Participants will be categorized into three groups:

1. 18–34 years
2. 35–50 years
3. 50+ years

## Data Analysis Method   
We will use **one-way ANOVA** to analyze the effects of pain reliever type on memory performance.
and dosage levels on cognitive retention.

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(janitor)
library(here)
library(lubridate)
library(arrow)
library(dplyr)
library(knitr)
library(performance)
library(kableExtra)
library(performance)

clean_data <- read.csv(here("data", "analysis_data", "clean_data.csv"))

```

# Summary Statistics

## Summary of quantitative variables

@tbl-summary-1 shows the summary statistics of game scores before and after treatment. The mean is similar across all groups, indicating no extreme changes in cognitive performance. However, the standard deviation (SD) is relatively high, suggesting some variation among participants.

- The median game scores remain close to the mean, showing a nearly symmetric distribution.
- The IQR (Interquartile Range) is smaller than the SD, meaning there may be outliers.
- The cards memory task shows little change before and after treatment, suggesting that pain relievers may not have a major effect on this specific task.

```{r}
#| echo: false
#| warning: false
#| message: false

clean_data$improvement <- clean_data$game_after - clean_data$game_before

improvement_summary <- clean_data %>%
  group_by(drug) %>%
  summarise(
    Mean_Improvement = mean(improvement, na.rm = TRUE),
    SD_Improvement = sd(improvement, na.rm = TRUE),
    Median_Improvement = median(improvement, na.rm = TRUE)
  )

kable(improvement_summary, format = "latex", booktabs = TRUE, caption = "Mean Improvement in Game Scores by Drug")

```

```{r}
#| label: tbl-summary-1
#| echo: false
#| warning: false
#| message: false

# References:
# https://tellingstorieswithdata.com/
# https://q.utoronto.ca/courses/376099/pages/project-descriptions?module_item_id=6379323

m1_game_before <- mean(clean_data$game_before, na.rm = TRUE)
m2_game_before <- median(clean_data$game_before, na.rm = TRUE)
m3_game_before <- sd(clean_data$game_before, na.rm = TRUE)
m4_game_before <- IQR(clean_data$game_before, na.rm = TRUE)

m1_game_after <- mean(clean_data$game_after, na.rm = TRUE)
m2_game_after <- median(clean_data$game_after, na.rm = TRUE)
m3_game_after <- sd(clean_data$game_after, na.rm = TRUE)
m4_game_after <- IQR(clean_data$game_after, na.rm = TRUE)

m1_cards_before <- mean(clean_data$cards_before, na.rm = TRUE)
m2_cards_before <- median(clean_data$cards_before, na.rm = TRUE)
m3_cards_before <- sd(clean_data$cards_before, na.rm = TRUE)
m4_cards_before <- IQR(clean_data$cards_before, na.rm = TRUE)

m1_cards_after <- mean(clean_data$cards_after, na.rm = TRUE)
m2_cards_after <- median(clean_data$cards_after, na.rm = TRUE)
m3_cards_after <- sd(clean_data$cards_after, na.rm = TRUE)
m4_cards_after <- IQR(clean_data$cards_after, na.rm = TRUE)

# Combine into a table
summary_table <- data.frame(
  Statistic = c("Mean", "Median", "SD", "IQR"),
  Game_Before = c(m1_game_before, m2_game_before, m3_game_before, m4_game_before),
  Game_After = c(m1_game_after, m2_game_after, m3_game_after, m4_game_after),
  Cards_Before = c(m1_cards_before, m2_cards_before, m3_cards_before, m4_cards_before),
  Cards_After = c(m1_cards_after, m2_cards_after, m3_cards_after, m4_cards_after)
)

kable(summary_table, format = "latex", booktabs = TRUE, caption = "Summary Statistics for Quantitative Variables") 
```

```{r}
#| echo: false
#| warning: false
#| message: false

# Count and percentage for Drug
drug_counts <- table(clean_data$drug)
drug_percentages <- prop.table(drug_counts) * 100
drug_summary <- data.frame(Drug = names(drug_counts), Count = as.numeric(drug_counts), Percentage = round(as.numeric(drug_percentages), 2))
colnames(drug_summary) <- c("Drug", "Count", "Percentage")

# Count and percentage for Dosage Level
dosage_counts <- table(clean_data$dosage_level)
dosage_percentages <- prop.table(dosage_counts) * 100
dosage_summary <- data.frame(Dosage = names(dosage_counts), Count = as.numeric(dosage_counts), Percentage = round(as.numeric(dosage_percentages), 2))

# Count and percentage for Location
location_counts <- table(clean_data$location)
location_percentages <- prop.table(location_counts) * 100
location_summary <- data.frame(Location = names(location_counts), Count = as.numeric(location_counts), Percentage = round(as.numeric(location_percentages), 2))

# Count and percentage for Age Group
age_counts <- table(clean_data$age_group)
age_percentages <- prop.table(age_counts) * 100
age_summary <- data.frame(Age_Group = names(age_counts), Count = as.numeric(age_counts), Percentage = round(as.numeric(age_percentages), 2))

# Print tables
kable(drug_summary, format = "latex", booktabs = TRUE, caption = "Summary of Drug Types")

kable(dosage_summary, format = "latex", booktabs = TRUE, caption = "Summary of Dosage Levels")

kable(location_summary, format = "latex", booktabs = TRUE, caption = "Summary of Locations")

kable(age_summary, format = "latex", booktabs = TRUE, caption = "Summary of Age Groups") 
```

\newpage

```{r}
#| echo: false
#| warning: false
#| message: false
#| fig.width: 9
#| fig.height: 8
#| dpi: 300
#| fig-cap: "Boxplot of Game Scores Before and After by Drug"

library(ggplot2)
library(patchwork)

#References
#https://r-graph-gallery.com/package/patchwork.html

clean_data$drug <- factor(clean_data$drug, labels = c("Aspirin", "Paracetamol", "Placebo", "Tramadol"))

# Create first boxplot
p1 <- ggplot(clean_data, aes(x = drug, y = game_before, fill = drug)) + 
  geom_boxplot() +
  labs(title = "Memory Game Scores Before Treatment", x = "Drug Type", y = "Game Score", fill = "Drug") +
  theme_minimal()

# Create second boxplot
p2 <- ggplot(clean_data, aes(x = drug, y = game_after, fill = drug)) + 
  geom_boxplot() +
  labs(title = "Memory Game Scores After Treatment", x = "Drug Type", y = "Game Score", fill = "Drug") +
  theme_minimal()

p1 / p2 + plot_annotation()

```

```{r}
#| echo: false
#| warning: false
#| message: false
#| fig.width: 9
#| fig.height: 8
#| dpi: 300
#| fig-cap: "Boxplot of Memory Card Scores Before and After by Drug"

library(ggplot2)
library(patchwork)

clean_data$drug <- factor(clean_data$drug, labels = c("Aspirin", "Paracetamol", "Placebo", "Tramadol"))

p1 <- ggplot(clean_data, aes(x = drug, y = cards_before, fill = drug)) + 
  geom_boxplot() +
  labs(title = "Memory Card Scores Before Treatment", x = "Drug Type", y = "Memory Card Score", fill = "Drug") +
  theme_minimal()

p2 <- ggplot(clean_data, aes(x = drug, y = cards_after, fill = drug)) + 
  geom_boxplot() +
  labs(title = "Memory Card Scores After Treatment", x = "Drug Type", y = "Memory Card Score", fill = "Drug") +
  theme_minimal()

p1 / p2 + plot_annotation()

```

```{r}
#| echo: false
#| warning: false
#| message: false
#| fig.width: 7
#| fig.height: 4
#| dpi: 300
#| fig-cap: "Distribution of Dosage Levels"

ggplot(clean_data, aes(x = dosage_level, fill = dosage_level)) +
  geom_bar() +
  labs(title = "Distribution of Dosage Levels", x = "Dosage Level", y = "Count") +
  theme_minimal()

```

```{r}
#| echo: false
#| warning: false
#| message: false
#| fig.width: 9
#| fig.height: 8
#| dpi: 300
#| fig-cap: "Assumption Check for Time taken after taking the Drug"

# Fit ANOVA Model
model <- aov(game_after ~ drug, data = clean_data)

# Check Model Assumptions
check_model(model, check = c("normality", "homogeneity"))
```

```{r}
#| include: false
#| warning: false
#| message: false
# Barlett Test Homegenity of Variance: testing for ass
# bartlett_test <- bartlett.test(game_after ~ drug, data = clean_data)
# print(bartlett_test)
```

```{r}
#| echo: false
#| warning: false
#| message: false
#| fig.width: 9
#| fig.height: 8
#| dpi: 300
#| fig-cap: "Assumption Check for Time taken before taking the Drug"

# Fit ANOVA Model
model_b <- aov(game_before ~ drug, data = clean_data)

# Check Model Assumptions
check_model(model_b, check = c("normality", "homogeneity"))
```

# References