### 2025-03-31 paper/paper.qmd
fixed plots and feedback fixes x2

---
title: "Effects of Different Pain Relievers and Dosages on Cognitive Retention"
subtitle: Results
author: 
  - Navya Hooda
  - Mohammed Yusuf Shaikh
  - Shivank Goel
  - Vanshika Vanshika
  - Jena Shah
date: today
date-format: long
abstract: Painkillers are commonly used to manage pain, and their effects are often associated with  physical relief. While they provide immediate physical comfort, their impact on cognitive performance is often overlooked. This study explores how different painkillers and dosages influence cognitive function. Using simulated testing from a virtual population, participants were randomly assigned a treatment (drug type and dosage) while controlling for variables such as age. Our findings indicate no significant relationship between drug type, dosage, and overall cognitive performance post-treatment. However, we observed some correlation between certain drug-dosage combinations and memory improvement, suggesting a potential area for further research. While the results were not statistically significant, this study contributes to a deeper understanding of the cognitive effects of painkillers and highlights the need for further investigation into their potential impact on memory and cognition.
number-sections: true
format: pdf
bibliography: references.bib
toc: true
---

# Acknowledgements

No auto-complete tools such as co-pilot were used in the course of this project, however, Language Learning Model ChatGPT was used while writing this paper. It was used for the purpose of code debugging, understanding models, and knowledge of certain topics, which we were not aware of. The chat with the AI bot is also attached as a reference under @sec-appendix

# Introduction

Pain can make everyday tasks harder, particularly when mental focus and clarity are required. Study has shown that individuals with pain tend to have cognitive dysfunction symptoms (Hu, 2010). Common pain relievers like Aspirin, Paracetamol, and Tramadol are often used to reduce physical discomfort, but their impact on cognitive functions are not well understood. Memory retention is a process that requires mental effort – which becomes more difficult when we're in pain. 
This study investigates how different pain relievers (Aspirin, Paracetamol, and Tramadol) at standard dosages affect memory performance. We will also look at how the interaction between drug type and dosage level influence memory function. Since the ability to recall information is a basis for optimized human function, understanding how pain relief might impact cognitive performance is important for managing pain without affecting productivity.

## Research Questions
1. *How does the type of pain reliever (Aspirin 500 mg, Paracetamol 500 mg, Tramadol 50 mg, and Placebo) affect memory game performance?*
- Null Hypothesis: All pain relievers at normal dosages act the same for cognitive retention game performances.
- Alternative Hypothesis: There exists at least one pair of distinct pain relievers at normal dosages which act differently for cognitive retention game performances.

2. *Is there an interaction effect between the type of pain reliever and dosage level on the improvement in memory game performance?*
- Null Hypothesis: There is no interaction effect between the type of pain reliever and dosage level on the improvement in memory task performance.
- Alternative Hypothesis: There is an interaction effect between at least one of the types of pain reliever and dosage level on the improvement in memory task performance.

# Methodology

In February, our group members collected observations from participants belonging to a virtual population to assess whether the treatment had any effect on memory performance, this resulted in a total of 240 observations. Each participant was **randomly** assigned a treatment.

The experiment utilized a two-way ANOVA design, where the experimental unit was the participant, the explanatory variables were the type of painkiller and dosage level (low/high), and the response variable was the memory performance, measured by the memory game score.

## Study Design and Treatments   
We selected three brands of painkillers: Aspirin, Paracetamol, and Tramadol. Additionally, we considered a placebo treatment. As a result, this study involves 2 factors: pain reliever type (4 levels - Aspirin, Paracetamol, Tramadol, Placebo) and dosage (2 levels - Low, High). Crossing the two factors provides us with a total of *8 treatment groups*, with **30 participants** assigned to each. The groups are as follows:

```{r, echo = FALSE}

library(knitr)
library(kableExtra)

# Create a dataframe for the table
study_table <- data.frame(
  Dosage = c("Low", "High"),
  Aspirin = c("500 mg", "1000 mg"),
  Paracetamol = c("500 mg", "1000 mg"),
  Tramadol = c("50 mg", "100 mg"),
  Placebo = c("Low", "High")
)

# Render the table
kable(study_table, caption = "Treatment Groups",
      align = "c") %>%
  kable_styling(full_width = FALSE, position = "center")
```

This results in **240 total observations** (8 × 30 = 240).

For each treatment, we selected 30 different people (n = 30). Through randomly picking the virtual citizens, we replicated 30 experimental units per treatment. The even observations per treatment group ensures we have a balanced design.

### Nuisance Variable: Age  
To reduce bias, we consider **controlling** the experiment by limiting participant ages to be strictly 18+. Hence in this study, **age** acts as a **nuisance variable**.
We also used **blocking** for age by ensuring participants were categorized into three even groups to prevent bias created through age. @tbl-summary-5 shows the breakdown of the groups.

## Quantitative and Qualitative Variables 
In our study the quantitative and qualitative variables are as follows:

1. Quantitative variables: Memory Game Scores, Memory Test Cards Scores 
2. Qualitative variables: Type of pain reliever, dosage level (low/high)

```{r}
#| include: false
#| warning: false
#| message: false
#| echo: false  # Hides code

library(tidyverse)
library(janitor)
library(here)
library(lubridate)
library(patchwork)
library(arrow)
library(dplyr)
library(knitr)
library(performance)
library(modelsummary)
library(kableExtra)
library(performance)
library(ggpubr)

clean_data <- read.csv(here("data", "analysis_data", "clean_data.csv"))

```

```{r}
#| label: tbl-summary-5
#| tbl-cap: "Summary of Age Groups"
#| echo: false
#| warning: false
#| message: false

age_counts <- table(clean_data$age_group)

# percentages for each age group
age_percentages <- prop.table(age_counts) * 100

# summary dataframe
age_summary <- data.frame(
  `Age Group` = names(age_counts),
  Count = as.numeric(age_counts),
  Percentage = round(as.numeric(age_percentages), 2)
)

colnames(age_summary)[colnames(age_summary) == "Age.Group"] <- "Age Group"

# Display table
kable(age_summary, booktabs = TRUE, row.names = FALSE)

```

The structure of the paper is as follows: Section 2 outlines our data and methodology. Section 3 presents our quantitative analysis. In Section 4, we discuss the results and address our findings from the analysis. Section 5 covers the limitations of our study, and Section 6 presents a conclusion to our analysis. After that, we have the bibliography in Section 7. Lastly, we attached the code for our study into the Appendix in Section 8.

## Summary of quantitative variables

@tbl-summary-1 shows the summary statistics of memory game scores before and after the treatment. The mean is similar across all groups, however, the standard deviation (SD) is relatively high. The median game scores remain close to the mean, and the IQR (Interquartile Range) is smaller than the SD. The cards memory task shows little change before and after treatment, suggesting that pain relievers may not have a major effect on this specific task.

Evident through the results, Paracetamol is the drug that has the greatest mean & median improvement in the game scores after administering the drug (without dosage taken into account). Aspirin appears to negatively impact the results of the game after being given to people, this was seen through the mean and median improvements.

```{r}
#| label: tbl-summary-1
#| tbl-cap: "Summary Statistics for Quantitative Variables"
#| echo: false
#| warning: false
#| message: false

# Compute statistics for Memory Game Score
mean_game <- mean(clean_data$memory_game_score, na.rm = TRUE)
median_game <- median(clean_data$memory_game_score, na.rm = TRUE)
sd_game <- sd(clean_data$memory_game_score, na.rm = TRUE)
iqr_game <- IQR(clean_data$memory_game_score, na.rm = TRUE)

# Compute statistics for Memory Cards Score
mean_cards <- mean(clean_data$memory_cards_score, na.rm = TRUE)
median_cards <- median(clean_data$memory_cards_score, na.rm = TRUE)
sd_cards <- sd(clean_data$memory_cards_score, na.rm = TRUE)
iqr_cards <- IQR(clean_data$memory_cards_score, na.rm = TRUE)

# Combine into a summary table
summary_table <- data.frame(
  Statistic = c("Mean", "Median", "SD", "IQR"),
  `Memory Game Score` = c(mean_game, median_game, sd_game, iqr_game),
  `Memory Cards Score` = c(mean_cards, median_cards, sd_cards, iqr_cards)
)

colnames(summary_table) <- c("Statistic", "Memory Game Score", "Memory Cards Score")

# Display table

kable(summary_table, format = "latex", booktabs = TRUE)

```

```{r}
#| label: tbl-summary-2
#| tbl-cap: "Summary of Drug Types"
#| echo: false
#| warning: false
#| message: false

drug_counts <- table(clean_data$drug)

# percentages
drug_percentages <- prop.table(drug_counts) * 100

#  summary dataframe
drug_summary <- data.frame(
  Drug = names(drug_counts),
  Count = as.numeric(drug_counts),
  Percentage = round(as.numeric(drug_percentages), 2)
)

# Display table
kable(drug_summary, format = "latex", booktabs = TRUE)

```

```{r}
#| label: tbl-summary-3
#| tbl-cap: "Summary of Dosage Levels"
#| echo: false
#| warning: false
#| message: false

dosage_counts <- table(clean_data$dosage_level)

# percentage for each dosage level
dosage_percentages <- prop.table(dosage_counts) * 100

# summary dataframe
dosage_summary <- data.frame(
  Dosage = names(dosage_counts),
  Count = as.numeric(dosage_counts),
  Percentage = round(as.numeric(dosage_percentages), 2)
)

# Display table
kable(dosage_summary, booktabs = TRUE, row.names = FALSE)

```

## Summary of counts and percentage for qualitative variables

@tbl-summary-2 shows the distribution of participants across drug types. Aspirin, Paracetamol, and Tramadol, and Placebo each have 60 participants (25\%), ensuring balance.

@tbl-summary-3 shows the dosage levels assigned to participants. There is an equal split between high-dose (120 participants, 50\%) and low-dose (120 participants, 50\%) conditions.

\newpage
# Plots

```{r, echo=FALSE}
#| label: fig-boxplot-game-scores
#| echo: false
#| warning: false
#| message: false
#| fig.width: 6
#| fig.height: 5
#| dpi: 300
#| out.width: 60%   # Adjust output width to 60% of the page width
#| out.height: 40%  # Adjust output height to 40% of the page height
#| fig-cap: "Boxplots: Memory Game Scores by Drug Type and Dosage Level"
#| layout-ncol: 2  # Ensures plots are side by side

p3 <- ggplot(clean_data, aes(x = drug, y = memory_game_score, fill = dosage_level)) +
  geom_boxplot() +
  labs(x = "Drug Type", y = "Memory Game Score", fill = "Dosage Level") +
  theme_minimal()
p3 
```

```{r}
#| label: fig-interaction
#| echo: false
#| warning: false
#| message: false
#| fig.width: 6  # Increased width
#| fig.height: 5  # Reduced height
#| dpi: 300
#| fig-cap: "Interaction Plot: Drug Type vs. Dosage Level"
#| layout-ncol: 2  # Ensures plots are side by side

library(ggplot2)
library(patchwork)  # For arranging plots

p1 <- ggplot(clean_data, aes(x = drug, y = memory_game_score, color = dosage_level, group = dosage_level)) +
  stat_summary(fun = mean, geom = "point", size = 2) +
  stat_summary(fun = mean, geom = "line", size = 1) +
  labs(title = "Memory Game Scores",
       x = "Drug Type",
       y = "Mean Memory Game Score",
       color = "Dosage Level") +
  theme_minimal()

# Arrange plots side by side with better spacing
p1

```
```{r}
#| label: fig-combined
#| echo: false
#| warning: false
#| message: false
#| fig.width: 12
#| fig.height: 6
#| fig-cap: "Interaction Plot and Boxplot: Memory Game Scores by Drug Type and Dosage Level"

# Combine the plots side by side
p1 + p3
```

## Game Score Distribution by Drug and Dosage Level

@fig-combined shows the distribution of memory game scores across different drug types using box plots.

1. The median scores for all drugs appear similar.
2. All drugs have a higher median on the higher dosage, level expect for Tramadol which is lower.
3. The Placebo group has a slightly lower median score compared to other drugs.
4. The (IQR) is **similar across all drug types**, though Tramadol and Placebo see a higher IQR in the memory game scores in the high dosage.

This boxplot provides an initial comparison of memory game performance and drug administration.

# Assumptions

## Model
Let Y denote the memory game score, then

$$
\text{Y}_i = \beta_0 + \beta_1 \cdot \text{drug}_i + \beta_2 \cdot \text{dosage\_level}_i + \beta_3 \cdot (\text{drug}_i \times \text{dosage\_level}_i) + \varepsilon_i
$$

Where:

- $\beta_0$ is the intercept.

- $\beta_1$, $\beta_2$, and $\beta_3$ are coefficients for the predictors and interaction term.

- $\varepsilon_i$ is the random error term.

```{r, echo = FALSE}
#| label: tbl-model-result
#| tbl-cap: "Regression Results for Memory Game Score"
#| echo: false
#| warning: false
#| message: false

# References: https://cran.r-project.org/web/packages/broom/vignettes/broom.html

library(broom)

# Regression Model results
anova_model <- lm(memory_game_score ~ drug * dosage_level, data = clean_data)

model_tidy <- tidy(anova_model)

kable(model_tidy, format = "latex", booktabs = TRUE)

```

```{r, echo = FALSE}
#| label: tbl-model-fit
#| tbl-cap: "Model Fit Statistics for Memory Game Score"
#| echo: false
#| warning: false
#| message: false

model_glance <- glance(anova_model)
kable(model_glance, format = "latex", booktabs = TRUE) %>%
   kable_styling(latex_options = "scale_down")
```

$$
Y_{ijk} = \mu + \alpha_i + \beta_j + (\alpha\beta)_{ij} + \varepsilon_{ijk}
$$

Where:

- $Y_{ijk}$: Memory score
- $\mu$: Overall mean  
- $\alpha_i$: Effect of the $i^{\text{th}}$ drug  
- $\beta_j$: Effect of the $j^{\text{th}}$ dosage level  
- $(\alpha\beta)_{ij}$: Interaction effect between drug and dosage  
- $\varepsilon_{ijk} \sim \mathcal{N}(0, \sigma^2)$: Random error

```{r, echo = FALSE}
#| label: tbl-anova-model
#| tbl-cap: "Two-Way ANOVA Results for Memory Game Score"
#| echo: false
#| warning: false
#| message: false

# Two way ANOVA result
anova_result <- anova(anova_model)

anova_table_df <- as.data.frame(anova_result)

kable(anova_table_df, format = "latex", booktabs = TRUE)

```

```{r, echo = FALSE}
#| label: fig-normal-plot
#| echo: false
#| warning: false
#| message: false
#| fig.width: 9  # Reduced width
#| fig.height: 8  # Reduced height
#| dpi: 300
#| out.width: 60%
#| out.height: 50%
#| fig.align: "center"
#| fig-cap: "Q-Q Plot of Residuals"

# Normality Check
ggqqplot(residuals(anova_model))

```

```{r, echo = FALSE}
#| label: tbl-shapiro
#| tbl-cap: "Shapiro-Wilk Test for Normality"
#| echo: false
#| warning: false
#| message: false

# Shapiro-Wilk Normality Test
shapiro_test <- shapiro.test(residuals(anova_model))
shapiro_table <- data.frame(
  Statistic = round(shapiro_test$statistic, 4),
  `P-Value` = round(shapiro_test$p.value, 4),
  Method = shapiro_test$method,
  check.names = FALSE  # Prevents automatic renaming
)

kable(shapiro_table, format = "latex", booktabs = TRUE)
```

```{r}
#| label: fig-variance-plot
#| echo: false
#| warning: false
#| message: false
#| fig.width: 9  # Smaller width
#| fig.height: 8  # Smaller height
#| out.width: 60%
#| out.height: 50%
#| dpi: 300
#| fig-cap: "Q-Q Plot of Residuals"

plot(anova_model, which = 3, main = "", sub = "", caption = "")

```

```{r}
#| label: tbl-bartlett
#| tbl-cap: "Bartlett’s Test for Homogeneity of Variance"
#| echo: false
#| warning: false
#| message: false

# Scale-Location Plot

# Perform Bartlett’s test for homogeneity of variance
bartlett_test <- bartlett.test(memory_game_score ~ interaction(drug, dosage_level), data = clean_data)

bartlett_table <- data.frame(
  Statistic = round(bartlett_test$statistic, 4),
  DF = bartlett_test$parameter,
  `P-Value` = round(bartlett_test$p.value, 4),
  Method = bartlett_test$method,
  check.names = FALSE
)

kable(bartlett_table, format = "latex", booktabs = TRUE)
```
# Analysis
The method of two-way ANOVA is used to estimate the mean of a memory game score changes according to the levels of two categorical variables, being drug type and dosage level. Our research question explores if there is an interaction effect between at least one of the types of pain reliever and dosage level on the improvement in memory task performance.

We verified the assumption of the two-way ANOVA model as seen from graph normality, using Q-Q plot as seen in @fig-normal-plot. The Shapiro-Wilk Test confirms a p-value = 0.05553 as per @tbl-shapiro, so we fail to reject the null hypothesis and the result confirms that the data-set is normally distributed.  
Our second assumption of Homogeneity of Variance is verified by the p-value = 0.3346 from the Bartlett test in @tbl-bartlett, which is greater than 0.05. Due to this, we fail to reject the null hypothesis.
Thirdly, for our two-way ANOVA model, the different combinations of pain reliever type and dosage levels were randomly assigned to all participants, in random order.

The two-way ANOVA factorial design involves the linear regression model followed by the ANOVA model. The model uses the reference (baseline) group which is Aspirin (Drug) & High (Dosage level). The regression model from the @tbl-model-result shows the intercept(beta_0) which represents the predicted mean of memory game score for participants receiving high dosage of aspirin is 62.20 and the p-value = 0.00000 indicates the intercept is statistically significant. The variance explained by the model is R^2 = 0.03983, approximately 4% of the response variable. And the p-value = 0.2167 of the model can be interpreted as not statistically significant

# Discussion
## Interpretation of Results  
The two-way ANOVA analysis revealed no statistically significant main effects of drug type (p = 0.25) or dosage level (p = 0.26) on memory game scores. Similarly, the interaction effect between drug type and dosage level was not significant (p = 0.24). These results suggest that neither the type of pain reliever (Aspirin, Paracetamol, Tramadol, or Placebo) nor its dosage (low vs. high) had a meaningful impact on memory game scores, which is representative of cognitive improvement in this study. The lack of significance persisted even after accounting for age as a confounding variable, which was controlled through participant stratification into three balanced groups (18–34, 35–50, and 50+ years).

# Limitations 
Our study is overall focused on memory, which may have links to one’s education and IQ level as they are all related to mental function and cognition. In the future, we could test everyone for IQ scores and their highest level of education – these two nuisance variables can be taken care of by using the four principles of experimental design. Additionally, we used 30 observations for each treatment group which may not be large enough. We can increase this number for future studies. To add on, not all participants were in pain, so the efficacy of the improvement of pain killers may be biased and using a virtual population may not ensure whether this aspect could be captured. Lastly, cognitive retention was measured immediately post-treatment – the long term impacts are not yet explored. To improve our study, we can try to  consider individuals with real pain symptoms before the study to ensure a reduction of bias, and provide a realistic test setting. Additionally, carrying the experiment out in the long-term would help with understanding the effects of painkillers and dosages on memory.

# Conclusion
This study found no conclusive evidence that pain relievers or their dosages significantly affect cognitive retention. However, the exploratory trends observed, particularly with Paracetamol and Tramadol, highlight the need for larger-scale studies to investigate subtle or context-dependent effects. Future research could incorporate longitudinal designs, broader age-specific analyses, and alternative cognitive metrics to deepen understanding of painkillers’ cognitive implications.

# References
(Ha, 2010) https://journals.lww.com/ejanaesthesiology/FullText/2010/02000/Amitriptyline_rather_than_lornoxicam_ameliorates.6.aspx

\newpage

# Appendix {#sec-appendix}

```{r}

library(tidyverse)
library(janitor)
library(here)
library(lubridate)
library(patchwork)
library(arrow)
library(dplyr)
library(knitr)
library(performance)
library(modelsummary)
library(kableExtra)
library(performance)
library(ggpubr)

clean_data <- read.csv(here("data", "analysis_data", "clean_data.csv"))
```

```{r}
# Compute statistics for Memory Game Score
mean_game <- mean(clean_data$memory_game_score, na.rm = TRUE)
median_game <- median(clean_data$memory_game_score, na.rm = TRUE)
sd_game <- sd(clean_data$memory_game_score, na.rm = TRUE)
iqr_game <- IQR(clean_data$memory_game_score, na.rm = TRUE)

# Compute statistics for Memory Cards Score
mean_cards <- mean(clean_data$memory_cards_score, na.rm = TRUE)
median_cards <- median(clean_data$memory_cards_score, na.rm = TRUE)
sd_cards <- sd(clean_data$memory_cards_score, na.rm = TRUE)
iqr_cards <- IQR(clean_data$memory_cards_score, na.rm = TRUE)

# Combine into a summary table
summary_table <- data.frame(
  Statistic = c("Mean", "Median", "SD", "IQR"),
  `Memory Game Score` = c(mean_game, median_game, sd_game, iqr_game),
  `Memory Cards Score` = c(mean_cards, median_cards, sd_cards, iqr_cards)
)

colnames(summary_table) <- c("Statistic", "Memory Game Score", "Memory Cards Score")

# Display table

kable(summary_table, format = "latex", booktabs = TRUE)
```

```{r}

drug_counts <- table(clean_data$drug)

# percentages
drug_percentages <- prop.table(drug_counts) * 100

#  summary dataframe
drug_summary <- data.frame(
  Drug = names(drug_counts),
  Count = as.numeric(drug_counts),
  Percentage = round(as.numeric(drug_percentages), 2)
)

# Display table
kable(drug_summary, format = "latex", booktabs = TRUE)

```

```{r}

dosage_counts <- table(clean_data$dosage_level)

# percentage for each dosage level
dosage_percentages <- prop.table(dosage_counts) * 100

# summary dataframe
dosage_summary <- data.frame(
  Dosage = names(dosage_counts),
  Count = as.numeric(dosage_counts),
  Percentage = round(as.numeric(dosage_percentages), 2)
)

# Display table
kable(dosage_summary, booktabs = TRUE, row.names = FALSE)

```

```{r}
age_counts <- table(clean_data$age_group)

# percentages for each age group
age_percentages <- prop.table(age_counts) * 100

# summary dataframe
age_summary <- data.frame(
  `Age Group` = names(age_counts),
  Count = as.numeric(age_counts),
  Percentage = round(as.numeric(age_percentages), 2)
)
colnames(age_summary)[colnames(age_summary) == "Age.Group"] <- "Age Group"

# Display table
kable(age_summary, booktabs = TRUE, row.names = FALSE)

```

```{r}

p1 <- ggplot(clean_data, aes(x = drug, y = memory_game_score, color = dosage_level, group = dosage_level)) +
  stat_summary(fun = mean, geom = "point", size = 3) +
  stat_summary(fun = mean, geom = "line", size = 1) +
  labs(title = "Memory Game Scores",
       x = "Drug Type",
       y = "Mean Memory Game Score",
       color = "Dosage Level") +
  theme_minimal()

# Interaction plot for Memory Cards Scores
p2 <- ggplot(clean_data, aes(x = drug, y = memory_cards_score, color = dosage_level, group = dosage_level)) +
  stat_summary(fun = mean, geom = "point", size = 3) +
  stat_summary(fun = mean, geom = "line", size = 1) +
  labs(title = "Memory Cards Scores",
       x = "Drug Type",
       y = "Mean Memory Cards Score",
       color = "Dosage Level") +
  theme_minimal()

# Combine the two plots vertically
combined_plot <- p1 / p2
combined_plot
```

```{r}

clean_data$dosage_level <- factor(clean_data$dosage_level, levels = c("Low", "High"))

# Boxplot of Memory Game Scores by Drug, colored by Dosage Level
ggplot(clean_data, aes(x = drug, y = memory_game_score, fill = dosage_level)) +
  geom_boxplot() +
  labs(
    x = "Drug Type",
    y = "Memory Game Score",
    fill = "Dosage Level"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 14, face = "bold"),
    axis.title = element_text(size = 12)
  )

```

```{r}
#| fig.width: 9
#| fig.height: 8
#| dpi: 300

ggplot(clean_data, aes(x = drug, y = memory_cards_score, fill = dosage_level)) +
  geom_boxplot() +
  labs(
    x = "Drug Type",
    y = "Memory Cards Score",
    fill = "Dosage Level"
  ) +
  theme_minimal()
```

```{r, echo = FALSE}

# References: https://cran.r-project.org/web/packages/broom/vignettes/broom.html

library(broom)

# Regression Model results
anova_model <- lm(memory_game_score ~ drug * dosage_level, data = clean_data)

model_tidy <- tidy(anova_model)

kable(model_tidy, format = "latex", booktabs = TRUE)

```

```{r}

model_glance <- glance(anova_model)
kable(model_glance, format = "latex", booktabs = TRUE) %>%
   kable_styling(latex_options = "scale_down")
```

```{r}
#| warning: false
#| message: false

# Two way ANOVA result
anova_result <- anova(anova_model)

anova_table_df <- as.data.frame(anova_result)

kable(anova_table_df, format = "latex", booktabs = TRUE)
```

```{r}
#| fig.width: 4
#| fig.height: 3
#| out.width: 40%
#| out.height: 40%
#| dpi: 300
#| fig-cap: "Q-Q Plot of Residuals"

# Normality Check
ggqqplot(residuals(anova_model))
```

```{r}

#| fig.width: 9
#| fig.height: 8
#| dpi: 300
#| fig-cap: "Q-Q Plot of Residuals"

# Shapiro-Wilk Normality Test
shapiro_test <- shapiro.test(residuals(anova_model))
shapiro_table <- data.frame(
  Statistic = round(shapiro_test$statistic, 4),
  P_Value = round(shapiro_test$p.value, 4),
  Method = shapiro_test$method
)

kable(shapiro_table, format = "latex", booktabs = TRUE)
```

```{r}

plot(anova_model, which = 3)
```