### 2025-04-06 paper/paper.qmd
File discussion done new updates

---
title: "Effects of Different Pain Relievers and Dosages on Cognitive Retention"
author: 
  - Navya Hooda
  - Mohammed Yusuf Shaikh
  - Shivank Goel
  - Vanshika Vanshika
  - Jena Shah
date: today
date-format: long
abstract: Painkillers are commonly used to manage pain and their effects are often associated with physical relief. While they provide immediate physical comfort, their impact on cognitive performance is often overlooked. This study explores how different painkillers and dosages influence cognitive function. Using simulated testing from a virtual population, participants were randomly assigned a treatment (drug type and dosage) while controlling for variables such as age. Our findings indicate no significant relationship between drug type, dosage, and overall cognitive performance post-treatment. However, we observed some correlation between certain drug-dosage combinations and memory improvement, suggesting a potential area for further research. While the results were not statistically significant, this study contributes to a deeper understanding of the cognitive effects of painkillers and highlights the need for further investigation into their potential impact on memory and cognition.
number-sections: true
format: pdf
bibliography: references.bib
toc: true
---

# Acknowledgements

No auto-complete tools such as co-pilot were used in the course of this project, however, the Language Learning Model 'ChatGPT' was used while writing this paper.
It was used for the purpose of code debugging, understanding models, and knowledge of certain topics, which we were not aware of.
The chat with the AI bot is also attached as a reference under @sec-appendix.

# Introduction {#sec-intro}

Pain can make everyday tasks harder, particularly when mental focus and clarity are required. Study has shown that individuals with pain tend to have cognitive dysfunction symptoms [@Hu]. Common pain relievers like Aspirin, Paracetamol, and Tramadol are often used to reduce physical discomfort, but their impact on cognitive functions are not well understood. Memory retention is a process that requires mental effort – which becomes more difficult when we are in pain. This study investigates how different pain relievers (Aspirin, Paracetamol, and Tramadol) at standard dosages affect memory performance. We will also take a look at how the interaction between drug type and dosage level influence memory function. Understanding how pain relief might impact cognitive performance is important since the ability to recall information is a basis for optimized human function.

## Research Questions

Our study consists of the following research questions.

Research Question 1.
*How does the type of pain reliever (Aspirin 500 mg, Paracetamol 500 mg, Tramadol 50 mg, and Placebo) affect memory game performance?*

-   Null Hypothesis: All pain relievers at normal dosages act the same for cognitive retention game performances.
-   Alternative Hypothesis: There exists at least one pair of distinct pain relievers at normal dosages which act differently for cognitive retention game performances.

Research Question 2.
*Is there an interaction effect between the type of pain reliever and dosage level on the improvement in memory game performance?*

-   Null Hypothesis: There is no interaction effect between the type of pain reliever and dosage level on the improvement in memory task performance.
-   Alternative Hypothesis: There is an interaction effect between at least one of the types of pain reliever and dosage level on the improvement in memory task performance.

The structure of this paper is as follows. @sec-method outlines our data and methodology. @sec-plots presents our findings into graphs and charts to visualize the data better. @sec-analysis presents our quantitative analysis. In @sec-disc, we discuss the results and address our findings. We end off, we have the limitations and conclusion in @sec-lims and @sec-conc, respectively.
All of the code used for this study can be found in @sec-appendix. The references are found in @sec-references.

# Methodology {#sec-method}

This study followed all four principles of experimental design: **Control, Block, Randomization and Replication**.

In February, our group members collected observations from participants belonging to a virtual population to assess whether the treatment had any effect on memory performance, this resulted in a total of 240 observations.
Each participant was **randomly** assigned a treatment.

The experiment utilized a two-way ANOVA design, where the experimental unit was the participant, the explanatory variables were the type of painkiller and dosage level (low/high), and the response variable was the memory performance, measured by the memory game score.

## Study Design and Treatments

We selected three brands of painkillers: Aspirin, Paracetamol, and Tramadol.
Additionally, we considered a placebo treatment.
As a result, this study involves 2 factors: pain reliever type (4 levels - Aspirin, Paracetamol, Tramadol, Placebo) and dosage (2 levels - Low, High).
Crossing the two factors provides us with a total of *8 treatment groups*, with **30 participants** assigned to each.
The total crossed treatment groups are shown in @tbl-total-groups.

This results in **240 total observations** (8 × 30 = 240).

For each treatment, we selected 30 different people (n = 30).
Through randomly picking the virtual citizens, we **replicated** 30 experimental units per treatment.
The even observations per treatment group ensures we have a **balanced design**.

```{r}
#| label: tbl-total-groups
#| tbl-cap: "Treatment Groups"
#| echo: false
#| warning: false
#| message: false

library(knitr)
library(kableExtra)

# Create a dataframe for the table
study_table <- data.frame(
  Dosage = c("Low", "High"),
  Aspirin = c("Aspirin 500 mg", "Aspirin 1000 mg"),
  Paracetamol = c("Paracetamol 500 mg", "Paracetamol 1000 mg"),
  Tramadol = c("Tramadol 50 mg", "Tramadol 100 mg"),
  Placebo = c("Placebo 1 Tablet", "Placebo 2 Tablets")
)

# Render the table
kable(study_table, align = "c") %>%
  add_header_above(c(" " = 1, "Drug Type" = 4)) %>%
  kable_styling(full_width = FALSE, position = "center")
```

### Nuisance Variable: Age

To reduce bias, we consider **controlling** the experiment by limiting participant ages to be strictly 18+.
Hence in this study, **age** acts as a **nuisance variable**.
We also used **blocking** for age by ensuring participants were categorized into three even groups to prevent bias created through age.
@tbl-summary-5 shows the breakdown of the groups.

## Quantitative and Qualitative Variables

In our study the quantitative and qualitative variables are as follows:

1.  Quantitative variables: Memory Game Scores
2.  Qualitative variables: Type of pain reliever, dosage level (low/high)

```{r}
#| include: false
#| warning: false
#| message: false
#| echo: false  # Hides code

library(tidyverse)
library(janitor)
library(here)
library(lubridate)
library(patchwork)
library(arrow)
library(dplyr)
library(knitr)
library(performance)
library(modelsummary)
library(kableExtra)
library(performance)
library(ggpubr)

clean_data <- read.csv(here("data", "analysis_data", "clean_data.csv"))

```

```{r}
#| label: tbl-summary-5
#| tbl-cap: "Summary of Age Groups"
#| echo: false
#| warning: false
#| message: false

age_counts <- table(clean_data$age_group)

# percentages for each age group
age_percentages <- prop.table(age_counts) * 100

# summary dataframe
age_summary <- data.frame(
  `Age Group` = names(age_counts),
  Count = as.numeric(age_counts),
  Percentage = round(as.numeric(age_percentages), 2)
)

colnames(age_summary)[colnames(age_summary) == "Age.Group"] <- "Age Group"

# Display table
kable(age_summary, booktabs = TRUE, row.names = FALSE)

```

## Summary of quantitative variables

@tbl-summary-1 shows the summary statistics of the treatment memory game scores. The mean memory game score is 62.58 seconds, with a median of 61.50 seconds, which indicates that the distribution is fairly symmetric.

```{r}
#| label: tbl-summary-1
#| tbl-cap: "Summary Statistics for Quantitative Variables"
#| echo: false
#| warning: false
#| message: false

mean_game <- mean(clean_data$memory_game_score, na.rm = TRUE)
median_game <- median(clean_data$memory_game_score, na.rm = TRUE)
sd_game <- sd(clean_data$memory_game_score, na.rm = TRUE)
iqr_game <- IQR(clean_data$memory_game_score, na.rm = TRUE)

# Create summary table with clean column header
summary_table <- data.frame(
  Statistic = c("Mean", "Median", "SD", "IQR"),
  "Memory Game Score" = c(mean_game, median_game, sd_game, iqr_game),  check.names = FALSE
)

# Display table
kable(summary_table, format = "latex", booktabs = TRUE)

```

```{r}
#| label: tbl-summary-2
#| tbl-cap: "Summary of Drug Types"
#| echo: false
#| warning: false
#| message: false

drug_counts <- table(clean_data$drug)

# percentages
drug_percentages <- prop.table(drug_counts) * 100

#  summary dataframe
drug_summary <- data.frame(
  Drug = names(drug_counts),
  Count = as.numeric(drug_counts),
  Percentage = round(as.numeric(drug_percentages), 2)
)

# Display table
kable(drug_summary, format = "latex", booktabs = TRUE)

```

```{r}
#| label: tbl-summary-3
#| tbl-cap: "Summary of Dosage Levels"
#| echo: false
#| warning: false
#| message: false

dosage_counts <- table(clean_data$dosage_level)

# percentage for each dosage level
dosage_percentages <- prop.table(dosage_counts) * 100

# summary dataframe
dosage_summary <- data.frame(
  Dosage = names(dosage_counts),
  Count = as.numeric(dosage_counts),
  Percentage = round(as.numeric(dosage_percentages), 2)
)

# Display table
kable(dosage_summary, booktabs = TRUE, row.names = FALSE)

```

## Summary of counts and percentage for qualitative variables

@tbl-summary-2 shows the distribution of participants across drug types.
Aspirin, Paracetamol, and Tramadol, and Placebo each have 60 participants (25%), ensuring balance.

@tbl-summary-3 shows the dosage levels assigned to participants.
There is an equal split between high-dose (120 participants, 50%) and low-dose (120 participants, 50%) conditions.

## Confounding Variable

@tbl-summary-5 shows the breakdown of age groups, which is a confounding variable in this study. The three age groups: 18–34, 35–50, and 50+ each have 80 participants (33.33\%).

\newpage

# Plots {#sec-plots}

```{r, echo=FALSE}
#| label: fig-boxplot-game-scores
#| echo: false
#| warning: false
#| message: false
#| fig.width: 6
#| fig.height: 5
#| dpi: 300
#| out.width: 60%   # Adjust output width to 60% of the page width
#| out.height: 40%  # Adjust output height to 40% of the page height
#| fig-cap: "Boxplots: Memory Game Scores by Drug Type and Dosage Level"
#| layout-ncol: 2  # Ensures plots are side by side

p3 <- ggplot(clean_data, aes(x = drug, y = memory_game_score, fill = dosage_level)) +
  geom_boxplot() +
  labs(x = "Drug Type", y = "Memory Game Score", fill = "Dosage Level") +
  theme_minimal()

```

```{r}
#| label: fig-interaction
#| echo: false
#| warning: false
#| message: false
#| fig.width: 6  # Increased width
#| fig.height: 5  # Reduced height
#| dpi: 300
#| fig-cap: "Interaction Plot: Drug Type vs. Dosage Level"
#| layout-ncol: 2  # Ensures plots are side by side

library(ggplot2)
library(patchwork)  # For arranging plots

p1 <- ggplot(clean_data, aes(x = drug, y = memory_game_score, color = dosage_level, group = dosage_level)) +
  stat_summary(fun = mean, geom = "point", size = 2) +
  stat_summary(fun = mean, geom = "line", size = 1) +
  labs(title = "Memory Game Scores",
       x = "Drug Type",
       y = "Mean Memory Game Score",
       color = "Dosage Level") +
  theme_minimal()
```

```{r}
#| label: fig-combined
#| echo: false
#| warning: false
#| message: false
#| fig.width: 12
#| fig.height: 6
#| fig-cap: "Interaction and Boxplots for Drug and Dosage Effects"

# Combine the plots side by side
p1 + p3
```

@fig-combined shows the distribution of memory game scores across different drug types using box plots and interaction plot.

From the box plots, we observe:

1. The median memory game scores are relatively similar across most drug types.
2. Tramadol (high dosage) shows a higher median score compared to other treatments.
3. The Placebo group has a slightly lower median score than the others.
4. The interquartile range (IQR) is consistent across all groups, although Tramadol (high dosage) has slightly more variability.

The interaction plot suggests an interaction effect between drug type and dosage level, as the lines are not parallel. The low-dosage line shows a sharp increase in memory scores for Tramadol and a dip for Placebo. However, the high-dosage line follows a more gradual trend. This indicates that the effect of dosage on memory performance may depend on the type of pain reliever administered.

Although the visual pattern indicates potential interaction, a formal Two-Way ANOVA is required to determine the statistical significance of this relationship.

# Analysis {#sec-analysis}

## Assumption Checks

The method of two-way ANOVA is used to estimate the mean of memory game score changes according to the levels of two categorical variables that are drug and dosage level. Before performing the Two-Way ANOVA, we verified the assumptions of the model as seen from using Q-Q plot as seen in @fig-normal-plot.

```{r}
#| echo: false
#| warning: false
#| message: false

anova_model <- lm(memory_game_score ~ drug * dosage_level, data = clean_data)

model_tidy <- tidy(anova_model) %>%
  rename(
    `Predictor Term` = term,
    `Estimate` = estimate,
    `Std. Error` = std.error,
    `t-statistic` = statistic,
    `p-value` = p.value
  )
```

```{r, echo = FALSE}
#| label: fig-normal-plot
#| echo: false
#| warning: false
#| message: false
#| fig.width: 9  # Reduced width
#| fig.height: 8  # Reduced height
#| dpi: 300
#| out.width: 60%
#| out.height: 50%
#| fig.align: "center"
#| fig-cap: "Q-Q Plot of Residuals"

# Normality Check
ggqqplot(residuals(anova_model))

```

The Shapiro-Wilk Test confirms a p-value = 0.05553 as per @tbl-shapiro, so we fail to reject the null hypothesis and the result confirms that the dataset is *normally distributed*.

```{r, echo = FALSE}
#| label: tbl-shapiro
#| tbl-cap: "Shapiro-Wilk Test for Normality"
#| echo: false
#| warning: false
#| message: false

# Shapiro-Wilk Normality Test
shapiro_test <- shapiro.test(residuals(anova_model))
shapiro_table <- data.frame(
  Statistic = round(shapiro_test$statistic, 4),
  `P-Value` = round(shapiro_test$p.value, 4),
  Method = shapiro_test$method,
  check.names = FALSE  # Prevents automatic renaming
)

kable(shapiro_table, format = "latex", booktabs = TRUE)
```

Our second assumption of Homogeneity of Variance appears to be satisfied. As shown in the @fig-residual-plot, the spread of the residual remains relatively consistent across the of fitted values. Further the result of the Bartlett test for homogeneity of variances from @tbl-bartlett, suggests the p-value is 0.3346. Since this p-value is greater than the 0.05 significance level, we fail to reject the null hypothesis of equal variances. Hence, the assumption is verified. 
Lastly, for our two-way ANOVA model, the different combinations of pain reliever type and dosage levels, the treatments were randomly assigned to all participants which ensures *independence*.

```{r}
#| label: tbl-bartlett
#| tbl-cap: "Bartlett’s Test for Homogeneity of Variance"
#| echo: false
#| warning: false
#| message: false

# Scale-Location Plot

# Perform Bartlett’s test for homogeneity of variance
bartlett_test <- bartlett.test(memory_game_score ~ interaction(drug, dosage_level), data = clean_data)

bartlett_table <- data.frame(
  Statistic = round(bartlett_test$statistic, 4),
  DF = bartlett_test$parameter,
  `P-Value` = round(bartlett_test$p.value, 4),
  Method = bartlett_test$method,
  check.names = FALSE
)

kable(bartlett_table, format = "latex", booktabs = TRUE)
```

```{r}
#| label: fig-residual-plot
#| echo: false
#| warning: false
#| message: false
#| fig.width: 9  
#| fig.height: 8  
#| dpi: 300
#| out.width: 60%
#| out.height: 50%
#| fig.align: "center"
#| fig-cap: "Residual plot: Check for Homoscedasticity"

plot(anova_model, which = 3, main = "", sub = "")
```

## Model

The method of Two-way ANOVA is used to estimate the mean of a memory game score changes according to the levels of two categorical variables that is drug and dosage level.
The general statistical form of the model is:

$$
Y_{ijk} = \mu + \alpha_i + \beta_j + (\alpha\beta)_{ij} + \varepsilon_{ijk}
$$

Where

-   $Y_{ijk}$: Memory score for the $k$-th participant receiving the $i$-th drug and $j$-th dosage level\
-   $\mu$: Overall mean\
-   $\alpha_i$: Effect of the $i$-th drug (Aspirin, Tramadol, Paracetamol, Placebo)
-   $\beta_j$: Effect of the $j$-th dosage level (High Dosage, Low dosage)\
-   $(\alpha\beta)_{ij}$: Interaction effect between drug and dosage\
-   $\epsilon_{ijk} \sim N(0, \sigma^2)$ : Random error

In this analysis where our focus in on Two-Way ANOVA Model and so we fit using linear regression model using dummy variable coding using Treatment Aspirin-High dosage as the baseline group.
The aim is to interpret estimated coefficients relative to reference group.
The model we used in our analysis can be expressed as:

\begin{align*}
Y_i =\ & \beta_0 
+ \beta_1 \cdot \text{Paracetamol}_i 
+ \beta_2 \cdot \text{Placebo}_i 
+ \beta_3 \cdot \text{Tramadol}_i \\
& + \beta_4 \cdot \text{LowDosage}_i 
+ \beta_5 \cdot (\text{Paracetamol}_i \times \text{LowDosage}_i) \\
& + \beta_6 \cdot (\text{Placebo}_i \times \text{LowDosage}_i) 
+ \beta_7 \cdot (\text{Tramadol}_i \times \text{LowDosage}_i)
+ \varepsilon_i
\end{align*}

Where

-   $\beta_0$: Intercept (Aspirin at High dosage — the reference group)\
-   $\beta_1, \beta_2, \beta_3$: Main effects of Paracetamol, Placebo, and Tramadol relative to Aspirin\
-   $\beta_4$: Effect of Low dosage relative to High\
-   $\beta_5, \beta_6, \beta_7$: Interaction effects for each drug under Low dosage\
-   $\epsilon_i$: Residual error for the $i$-th participant

The Intercept term $\beta_0$ accounts for the reference group.
All other coefficients $\beta_j$ represent the mean difference between the specified group and the reference group, holding other factors constant for j = *1, 2, 3, 4, 5, 6, 7*

The @tbl-model-result show the intercept ($\beta_0$ = 62.20) is the average memory game score for individuals given High dosage and Aspirin Drug (reference group). While comparing other treatments to reference level, we observe most coefficients from @tbl-model-result  are not statistically significant at 5% level. The coefficient of Drug Paracetamol ($p-value = 0.9927$), Tramadol ($p-value = 0.5281$), Placebo ($p-value = 0.3893$) indicates that mean memory game scores for these groups are not statistically different from the reference group. Similarly, the evidence for all the Interaction terms suggests no statistical significance between drug type and dosage level has any effect on the memory game score performance as p value are greater than 5% significance level. The lack of significance indicates that the effect of dosage does not vary meaningfully across drug types and dosage levels, indicating an absence of interaction between the two factors.

```{r, echo = FALSE}
#| label: tbl-model-result
#| tbl-cap: "Coefficient Estimates from Linear Model for Memory Game Score"
#| echo: false
#| warning: false
#| message: false

# References: https://cran.r-project.org/web/packages/broom/vignettes/broom.html

library(broom)

anova_model <- lm(memory_game_score ~ drug * dosage_level, data = clean_data)

model_tidy <- tidy(anova_model) %>%
  rename(
    `Predictor Term` = term,
    `Estimate` = estimate,
    `Std. Error` = std.error,
    `t-statistic` = statistic,
    `p-value` = p.value
  )

kable(model_tidy, format = "latex", booktabs = TRUE)
```

```{r, echo = FALSE}
#| label: tbl-model-fit
#| tbl-cap: "Model Summary Statistics for Memory Game Score"
#| echo: false
#| warning: false
#| message: false

model_glance <- glance(anova_model) %>%
  transmute(
    `R - sqaured` = r.squared,
    `F-Statistic` = statistic,
    `Model df` = df,
    `Residual df` = df.residual,
    `p-value` = p.value,
    `Residual Std. Error` = sigma,
    `Sample Size (n)` = nobs
  )

kable(model_glance, format = "latex", booktabs = TRUE) %>%
  kable_styling(latex_options = "scale_down")

```

```{r, echo = FALSE}
#| label: tbl-anova-model
#| tbl-cap: "Two-Way ANOVA Results for Memory Game Score"
#| echo: false
#| warning: false
#| message: false

# Two way ANOVA result
anova_result <- anova(anova_model)

anova_table_df <- as.data.frame(anova_result)

kable(anova_table_df, format = "latex", booktabs = TRUE)

```

# Discussion {#sec-disc}

## Interpretation of Results

The two-way ANOVA results @tbl-anova-model revealed no statistically significant main effects of drug type ($p = 0.25$) or dosage level ($p = 0.26$) on memory game scores, as both p-values exceed the 0.05 threshold.

For **Research Question 1**, which asked whether the type of pain reliever impacts memory performance, we found no significant differences in memory scores between the drug groups (Aspirin, Paracetamol, Tramadol, or Placebo). Therefore, we fail to reject the null hypothesis for Research Question 1.

Similarly, the **interaction effect** between drug type and dosage level was not statistically significant ($p = 0.24$), suggesting that the impact of drug type on memory retention does not vary by dosage level. This directly addresses **Research Question 2**, and we again fail to reject the null hypothesis, concluding that there is no interaction effect between the type of pain reliever and dosage on memory task performance.

Since no interaction effect was present, we examined the main effects independently. As mentioned, drug type was not significant, and neither was **dosage level** ($p = 0.26$), indicating that neither factor significantly influenced cognitive performance.

The lack of statistically significant findings persisted even after **controlling for age** as a potential confounding variable. Participants were stratified into balanced age groups (18–34, 35–50, and 50+ years) to minimize age-related bias.

## Effect Size and Power Analysis

Since the interaction effect was not statistically significant ($p > 0.05$), we proceeded to assess the effect size with the purpose to evaluate the practical significance of our findings. Our model yields  $R^2 = 0.0398$. While $R^2$ commonly used in regression analysis to a measure explained variance, determining the goodness of fit of a model. However, in ANOVA Model the $R^2$ value doesn't have a direct interpretation as the proportion of variance explained by regression model and it also does not directly reflect differences between group means. We use the $R^2$ value to report the effect that corresponds to Cohen’s f guidelines i.e. $| f |$ is approximately 0.25 as our  $R^2$ falls between the small-medium categories.Even though the statistical test did not yield significant results, this quantity suggests we have a small-medium effect size, indicating that the type of drug and dosage level may still have had a meaningful impact on memory task performance.

We did a post-hoc power analysis. Based on the overall $\eta^2$ from the ANOVA model, we calculated a Cohen's $f$ value of 0.204. This suggests a small-medium effect size, which means that the type of drug and dosage level may still have had a meaningful impact on memory task performance.

The estimated power of the study was around 59.6%. This means that our study was moderately able to detect an effect with the size we have, if there was an effect. Despite being less than the ideal 80% power value, this suggests that the study is still able to detect large & meaningful differences -- just not smaller ones.

# Limitations {#sec-lims}

Our study is overall focused on memory, which may have links to one’s education and IQ level, as they are all related to mental function and cognition. In the future, we could test everyone for IQ scores and their highest level of education – these two nuisance variables can be taken care of by using the four principles of experimental design.
Additionally, we used 30 observations for each treatment group which may not be large enough. We can increase this number for future studies. To add on, not all participants were in pain, so the efficacy of the improvement of pain killers may be biased and using a virtual population may not ensure whether this aspect could be captured.
Lastly, cognitive retention was measured immediately post-treatment – the long term impacts are not yet explored. To improve our study, we can try to consider individuals with real pain symptoms before the study to ensure a reduction of bias, and provide a realistic test setting. Additionally, carrying the experiment out in the long-term would help with understanding the effects of painkillers and dosages on memory.

There’s also a chance that the participants’ expectations influenced their results if they knew which treatment they were getting (we cannot expand on this as it was a virtual experiment). In the future, bias could be minimized through having a double-blind design in which neither the participants nor the researchers know who’s getting what. We only tested low and high dosages, so we might be missing more subtle effects that show up with intermediate doses. So if we were to include a wider range of dosages in future studies, it could give a fuller picture. Another issue is that our sample might not fully represent the general population since we selected participants from specific clusters of locations -- not all locations were covered equally. This means our findings might not apply to everyone. Next time, we could choose more diverse and evenly spread-out clusters to make the results more general. Lastly, despite our efforts to control for confounding factors, there are still some variables (like sleep quality, stress levels, or baseline cognitive function) that could affect memory performance. Being more thorough in accounting for these factors could improve the accuracy of the results.

# Conclusion {#sec-conc}

This study did not find clear evidence that pain relievers or their dosages have a significant impact on cognitive retention. However, the exploratory trends observed, particularly with Paracetamol and Tramadol, highlight the need for larger-scale studies to investigate subtle or context-dependent effects. Future research could incorporate longitudinal designs, broader age-specific analyses, and alternative cognitive metrics to deepen understanding of painkillers’ cognitive implications.

\newpage

# Appendix {#sec-appendix}

```{r}

library(knitr)
library(kableExtra)

# Create a dataframe for the table
study_table <- data.frame(
  Dosage = c("Low", "High"),
  Aspirin = c("Aspirin 500 mg", "Aspirin 1000 mg"),
  Paracetamol = c("Paracetamol 500 mg", "Paracetamol 1000 mg"),
  Tramadol = c("Tramadol 50 mg", "Tramadol 100 mg"),
  Placebo = c("Placebo 1 Tablet", "Placebo 2 Tablets")
)

# Render the table
kable(study_table, align = "c") %>%
  add_header_above(c(" " = 1, "Drug Type" = 4)) %>%
  kable_styling(full_width = FALSE, position = "center")
```

```{r}

library(tidyverse)
library(janitor)
library(here)
library(lubridate)
library(patchwork)
library(arrow)
library(effectsize)
library(dplyr)
library(knitr)
library(performance)
library(modelsummary)
library(kableExtra)
library(performance)
library(WebPower)
library(ggpubr)

clean_data <- read.csv(here("data", "analysis_data", "clean_data.csv"))
```

```{r}
age_counts <- table(clean_data$age_group)

# percentages for each age group
age_percentages <- prop.table(age_counts) * 100

# summary dataframe
age_summary <- data.frame(
  `Age Group` = names(age_counts),
  Count = as.numeric(age_counts),
  Percentage = round(as.numeric(age_percentages), 2)
)

colnames(age_summary)[colnames(age_summary) == "Age.Group"] <- "Age Group"

# Display table
kable(age_summary, booktabs = TRUE, row.names = FALSE)

```

```{r}
mean_game <- mean(clean_data$memory_game_score, na.rm = TRUE)
median_game <- median(clean_data$memory_game_score, na.rm = TRUE)
sd_game <- sd(clean_data$memory_game_score, na.rm = TRUE)
iqr_game <- IQR(clean_data$memory_game_score, na.rm = TRUE)

# Create summary table with clean column header
summary_table <- data.frame(
  Statistic = c("Mean", "Median", "SD", "IQR"),
  "Memory Game Score" = c(mean_game, median_game, sd_game, iqr_game),  check.names = FALSE
)

# Display table
kable(summary_table, format = "latex", booktabs = TRUE)

```

```{r}

drug_counts <- table(clean_data$drug)

# percentages
drug_percentages <- prop.table(drug_counts) * 100

#  summary dataframe
drug_summary <- data.frame(
  Drug = names(drug_counts),
  Count = as.numeric(drug_counts),
  Percentage = round(as.numeric(drug_percentages), 2)
)

# Display table
kable(drug_summary, format = "latex", booktabs = TRUE)
```

```{r}
p3 <- ggplot(clean_data, aes(x = drug, y = memory_game_score, fill = dosage_level)) +
  geom_boxplot() +
  labs(x = "Drug Type", y = "Memory Game Score", fill = "Dosage Level") +
  theme_minimal()
```

```{r}
p1 <- ggplot(clean_data, aes(x = drug, y = memory_game_score, color = dosage_level, group = dosage_level)) +
  stat_summary(fun = mean, geom = "point", size = 2) +
  stat_summary(fun = mean, geom = "line", size = 1) +
  labs(title = "Memory Game Scores",
       x = "Drug Type",
       y = "Mean Memory Game Score",
       color = "Dosage Level") +
  theme_minimal()
```

```{r}
# Combine the plots side by side
p1 + p3
```

```{r}

clean_data$dosage_level <- factor(clean_data$dosage_level, levels = c("Low", "High"))

# Boxplot of Memory Game Scores by Drug, colored by Dosage Level
ggplot(clean_data, aes(x = drug, y = memory_game_score, fill = dosage_level)) +
  geom_boxplot() +
  labs(
    x = "Drug Type",
    y = "Memory Game Score",
    fill = "Dosage Level"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 14, face = "bold"),
    axis.title = element_text(size = 12)
  )

```

```{r}
#| fig.width: 9
#| fig.height: 8
#| dpi: 300

ggplot(clean_data, aes(x = drug, y = memory_cards_score, fill = dosage_level)) +
  geom_boxplot() +
  labs(
    x = "Drug Type",
    y = "Memory Cards Score",
    fill = "Dosage Level"
  ) +
  theme_minimal()
```

```{r, echo = FALSE}

# References: https://cran.r-project.org/web/packages/broom/vignettes/broom.html

library(broom)

# Regression Model results
anova_model <- lm(memory_game_score ~ drug * dosage_level, data = clean_data)

model_tidy <- tidy(anova_model)

kable(model_tidy, format = "latex", booktabs = TRUE)

```

```{r}

model_glance <- glance(anova_model)
kable(model_glance, format = "latex", booktabs = TRUE) %>%
   kable_styling(latex_options = "scale_down")
```

```{r}
#| warning: false
#| message: false

# Two way ANOVA result
anova_result <- anova(anova_model)

anova_table_df <- as.data.frame(anova_result)

kable(anova_table_df, format = "latex", booktabs = TRUE)
```

```{r}
#| fig.width: 4
#| fig.height: 3
#| out.width: 40%
#| out.height: 40%
#| dpi: 300
#| fig-cap: "Q-Q Plot of Residuals"

# Normality Check
ggqqplot(residuals(anova_model))
```

```{r}

#| fig.width: 9
#| fig.height: 8
#| dpi: 300
#| fig-cap: "Q-Q Plot of Residuals"

# Shapiro-Wilk Normality Test
shapiro_test <- shapiro.test(residuals(anova_model))
shapiro_table <- data.frame(
  Statistic = round(shapiro_test$statistic, 4),
  P_Value = round(shapiro_test$p.value, 4),
  Method = shapiro_test$method
)

kable(shapiro_table, format = "latex", booktabs = TRUE)
```

```{r}

plot(anova_model, which = 3)
```

```{r}
# EFFECT SIZE AND POWER ANALYSIS
eta_sq <- eta_squared(anova_model, partial = FALSE)

etasquare <- sum(eta_sq$Eta2)  # sum of all eta^2 values
cohens_f <- sqrt(etasquare / (1 - etasquare))

power_result <- wp.kanova(
  n = 240,
  ndf = 7, # because 8 - 1 = 7
  f = cohens_f,
  ng = 8, # because 8 treatment groups
  alpha = 0.05
)

```

\newpage

# References {#sec-references}